---
title: "GEMS_SMURP"
format: html
editor: visual
---

GEMS SMURP N15 label tests.

```{r}
library(tidyverse)
source("R/gems_functions.R")
```

## Nalgene Control

2023-11-16

Control Nalgene with sediment + seawater, no label. DAQ using PC and old RGASoft software.

```{r}
df <- read_rgasoft("data/RGASoft/GEMS_test_NF1_SMURP_control_2023-11-17.txt")
plot_gems(df)
```

## Nalgene Label

2023-11-18

First Nalgene label experiment. Nalgene with sediment + seawater, with label. DAQ using PC and old RGASoft software.

```{r}
df <- read_rgasoft("data/RGASoft/GEMS_test_NF1_SMURP_label_2023-11-21.txt")
plot_gems(df)
```

## System test

2023-11-27

Testing Teensy DAQ. Background, mostly water pump off.

```{r}
df <- read_gems("data/GEMS_tests/SerialCapture/Serial_capture_bench_test2023-11-27.txt")
plot_gems(df)
```

Lander SD data test. 2023-12-05. Water pump off.

```{r}
df <- read_gems("data/GEMS_tests/LanderSD/23_12_05_10_56_00.txt")
plot_gems(df)
```

## Open Nalgene

2023-12-08

Top of Nalgene open. Using fresh sediment from Wood Neck. Very stinky and fine grained.

```{r}
df <- read_gems("data/GEMS_tests/SerialCapture/GEMS_test_NF0_SMURP_label_open_2023-12-08.txt")
plot_gems(df)
```

## Closed Nalgene

2023-12-11

Repeat of previous experiment with nalgene sealed. Same Wood Neck sediment.
Power issue during run required reboot of system. Scroll pump failed when
power lost, swapped for spare. System time breifly off.

```{r}
df <- read_gems("data/GEMS_tests/SerialCapture/GEMS_test_NF0_SMURP_label_closed_2023-12-11.txt") |> 
  filter(timestamp > "2023-12-01")
plot_gems(df)
```