---
title: "GEMS Nalgene SMURP Tests"
format: html
self-contained: true
editor: source
execute:
  message: false
  warning: false
---

GEMS SMURP N15 label tests and background runs Nov-Dec 2023.

```{r}
library(tidyverse)
library(dygraphs)
source("R/gems_functions.R")
theme_set(theme_classic())
```

## First Nalgene Experiment

This experiment was run using a PC to control the RGA and DAQ using old RGASoft software. 
Pressure values are different from the data when using Teensy for DAQ. Likely just an issue when converting from current to pressure.

```{r}
rga_file <- "data/RGASoft/GEMS_test_NF1_SMURP_label_2023-11-21.txt"
control_start <-  as.POSIXct("2023-11-16 14:22:00", tz="UTC")
label_start <-  as.POSIXct("2023-11-17 11:22:00", tz="UTC")
df <- read_rgasoft(rga_file) |> 
  mutate(experiment = case_when(timestamp < control_start ~ "warmup",
                                timestamp < label_start ~ "control",
                                TRUE ~ "label"))

#write_csv(df, "data/GEMS_SMURP/nalgene_SMURP_2023-11-17.csv")
```

### Warmup

2023-11-16

Warmup and pumpdown. High values are pump on circulating freshwater from beaker. Low values are pump off.

```{r}
df |> 
  filter(experiment == "warmup") |> 
  plot_gems()
```
### Control

2023-11-16

Control Nalgene with sediment + seawater, no label. 

```{r}
df |> 
  filter(experiment == "control") |> 
  plot_gems()
```

### Nalgene Label

2023-11-18

First Nalgene label experiment. Nalgene with sediment + seawater, with label. 

```{r}
df |> 
  filter(experiment == "label") |> 
  plot_gems()
```

## System test

2023-11-27

Testing Teensy DAQ. Background, mostly water pump off.

```{r}
df <- read_gems("data/GEMS_tests/SerialCapture/Serial_capture_bench_test2023-11-27.txt")
plot_gems(df)
```

Lander SD data test. 2023-12-05. Water pump off.

```{r}
df <- read_gems("data/GEMS_tests/LanderSD/23_12_05_10_56_00.txt")
plot_gems(df)
```

## Open Nalgene

2023-12-08

Top of Nalgene open. Using fresh sediment from Wood Neck. Very stinky and fine grained.

```{r}
df <- read_gems("data/GEMS_tests/SerialCapture/GEMS_test_NF0_SMURP_label_open_2023-12-08.txt")
plot_gems(df)
```

## Closed Nalgene

2023-12-11

Repeat of previous experiment with nalgene sealed. Same Wood Neck sediment. Power issue during run required reboot of system. Scroll pump failed when power lost, swapped for spare. System time breifly off.

```{r}
df <- read_gems("data/GEMS_tests/SerialCapture/GEMS_test_NF0_SMURP_label_closed_2023-12-11.txt") |> 
  filter(timestamp > "2023-12-01")
plot_gems(df)
```

# Full SMURP experiments

## 2024-04-26

First successful full SMURP experiment. 
SMURP transferred to large black crate.
Added XX l of seawater from ESL to bring level to ~3cm from box top.
Pump installed 
Inlet filtered with copper LECS-style mesh
aquarium stone on outlet.

Initial data is background.
Step on April 23 was due to aquarium stone on outlet falling off. 
Pressure differential caused increased outgassing.

```{r}
label_start <-  as.POSIXct("2024-04-22 16:40:00")
no_airstone_start <-  as.POSIXct("2024-04-23 13:33:40")

df <- read_gems("data/GEMS_tests/SerialCapture/GEMS_full_SMURP_NF0_2024-04-22.txt") |> 
  mutate(experiment = case_when(timestamp < label_start ~ "control",
                                timestamp < no_airstone_start ~ "label-airstone",
                                TRUE ~ "label-no-airstone"))

df |> 
  filter(mass == 32) |> 
  select(timestamp, pressure) |> 
  dygraphs::dygraph()
plot_gems(df)

df |> 
  write_csv("data/GEMS_tests/SerialCapture/GEMS_full_SMURP_NF0_proc_2024-04-22.csv")

```

```{r}
df |> 
  filter(mass %in% c(29, 30)) |> 
  plot_gems()
```

## 2024-04-30

Closed sided sediment only SMURP control.

Added XX l of seawater from ESL to bring level to ~3cm from box top.
Pump installed 
Inlet filtered with copper LECS-style mesh

Initial data is background.

```{r}
label_start <-  as.POSIXct("2024-04-22 16:40:00")

df <- read_gems("data/GEMS_tests/SerialCapture/GEMS_full_SMURP_sed_only_NF0_2024-04-30.txt") |> 
  mutate(experiment = case_when(timestamp < label_start ~ "control",
                                TRUE ~ "label"))

plot_gems(df)

df |> 
  write_csv("data/GEMS_tests/SerialCapture/GEMS_full_SMURP_sed_only_NF0_proc_2024-04-30.csv")
```


```{r}
df |> 
  filter(mass %in% c(29, 30)) |> 
  pivot_wider(names_from = mass, values_from = pressure, names_prefix = "mass_") |> 
  select(timestamp, mass_29, mass_30) |> 
  dygraphs::dygraph()
```

```{r}
df |> 
  filter(mass %in% c(29, 30)) |> 
  plot_gems()
```

### Rate calculations

```{r}
df_proc <- process_gems()
```


Firesting data

```{r}
read_firesting_go2 <- function(file) {
  cn <- read_lines(file, skip = 15, n_max = 1, 
                   locale = locale(encoding = 'latin1')) |> 
    str_split("\t") |>
    unlist()
  df <- read_tsv(file, skip = 16)
  names(df) <- cn
  df
}
```

```{r}
files <- list.files("data/Firesting", full.names = TRUE)
sed_files <- files[4:5]
df <- map(sed_files, read_firesting_go2) |> 
  bind_rows()
```

```{r}
ggplot(df, aes(`Date_Time(YYYY-MM-DD_hh:mm:ss)`, `Oxygen(%air sat.)`)) + 
  geom_point()
```

```{r}
ggplot(df, aes(`Date_Time(YYYY-MM-DD_hh:mm:ss)`, `Temperature(°C)`)) + 
  geom_point()
```

```{r}
write_csv(df, "data/firesting_o2_smurp_sed_only_proc_2024-04-30.csv")
```


## 2024-05-13

Open sided sediment only SMURP control.

Added XX l of seawater from ESL to bring level to ~3cm from box top.
Pump installed 
Inlet filtered with copper LECS-style mesh

Initial data is background.

Not much mass 30 response. O2 at 20% when RGA stopped on Friday.
Left box in sink over weekend, O2 at 0% monday morning.

```{r}
label_start <-  as.POSIXct("2024-05-13 10:40:00")

df <- read_gems("data/GEMS_tests/SerialCapture/GEMS_full_SMURP_sed_crate_NF0_2024-05-13.txt") |> 
  mutate(experiment = case_when(timestamp < label_start ~ "control",
                                TRUE ~ "label"))

plot_gems(df)

df |> 
  write_csv("data/GEMS_tests/SerialCapture/GEMS_full_SMURP_sed_crate_NF0_proc_2024-05-13.csv")
```

```{r}
read_firesting <- function(file) {
  cn <- read_lines(file, skip = 10, n_max = 1, 
                   locale = locale(encoding = 'latin1')) |> 
    str_split("\t") |>
    unlist()
  df <- read_tsv(file, skip = 11)
  names(df) <- cn
  df
}
```

```{r}
files <- c("data/SMURP_O2_sed_crate_2024-05-13.txt", "data/SMURP_O2_sed_crate_2024-05-15.txt")
df <- map(files, read_firesting) |> 
  list_rbind()

write_csv(df, "data/SMURP_O2_sed_crate_proc_2024-05-13.txt")
```

```{r}
ggplot(df, aes(`DateTime (YYYY-MM-DD hh:mm:ss)`, `Oxygen (%air sat.)`)) + 
  geom_line()
```

```{r}
dygraph(select(df, `DateTime (YYYY-MM-DD hh:mm:ss)`, `Oxygen (%air sat.)`))
```

## 2024-08-19 First chamber experiment

Chamber label experiment in sed-only, closed sided crate.

* Installed chamber. Sed-surface 10cm, chamber-surface 5cm.
* Started pump, purge bubbles. Large initial bubble below screen.
* Inject 5ml of argon-saturated 20 mmol/L 15NO3- solution.
* Power outage

```{r}
label_start <-  as.POSIXct("2024-08-19 13:40:00")

df <- read_gems("data/GEMS_tests/SerialCapture/GEMS_2024-08-19_chamber_sed_closed.txt") |> 
  mutate(timestamp = ifelse(timestamp > "2024-08-18",
                            timestamp,
                            timestamp + 177775450),
         timestamp = as.POSIXct(timestamp),
         experiment = case_when(timestamp < label_start ~ "control",
                                TRUE ~ "label"))

plot_gems(df)

df |> 
  write_csv("data/GEMS_tests/SerialCapture/GEMS_2024-08-19_chamber_sed_closed_proc.csv")
```

Calculate slope and rate using Scott's method.

Give each measurement loop an average timestamp and pivot to wide format.


```{r}
df_wide <- rga_wider(df)
```

Natural nitrogen abundance and saturation concentration

```{r}
frac_abd_15 <- 0.00367
frac_abd_14 <- 1 - frac_abd_15
nit_sat_umol <- 600 # calculate for lab temp, sal, and pressure

```

Plot to find rough background values (should do this programmatically)

```{r}
df_wide %>% 
  select(timestamp, mass_28, mass_29, mass_30, mass_40) %>% 
  dygraph() %>% 
  dyOptions(logscale = TRUE)
```

Backgrounds from plot

```{r}
bg_29 <- 7.4E-12
bg_30 <- 2.1E-12
```

Normalized partial pressures and molar concentrations

```{r}
df_proc <- norm_rga(df_wide, bg_29, bg_30, nit_sat_umol = 600)
```

Nitrogen normalized argon. Use slope to calculate diffusive loss.

```{r}
df_proc %>% 
  select(et, mass_40_28) %>% 
  mutate(et = as.integer(et)) %>% 
  dygraph() %>% 
  dyOptions(logscale = TRUE)
```

```{r}
ar <- df_proc %>% 
  select(et, mass_40_28) %>% 
  filter(et > 10000,
         et < 20000)
```

```{r}
ar_loss_rate <- coef(lm(mass_40_28 ~ et, data = ar))[2] * 3600 * 24
```

pick time with even slope

```{r}
df_proc %>% 
  select(et, umol_30) %>% 
  ggplot(aes(et, umol_30)) +
  geom_line()
  mutate(et = as.integer(et)) %>% 
  dygraph() %>% 
  dy
%>% 
  dyOptions(logscale = TRUE)
```

Use least-squares linear regression to get slope of mass 30 production

```{r}
sdf <- df_proc %>% 
  select(et, umol_30) %>% 
  filter(et > 10000,
         et < 30000)
```

Slope (umol/s) of linear fit for inital linear range. Convert to umol/day

```{r}
umol_30_rate <- coef(lm(umol_30 ~ et, data = sdf))[2] * 3600 * 24
```

Initial rate of mass 30 production: `umol_30_rate` umol/day.

Rolling slope for instantaneous rate

```{r}
library(zoo)

df_slope <- df_proc %>% 
  select(et, umol_30) %>% 
  mutate(et = as.integer(et)) %>% 
  filter(et > 10000)

lm_slope <- function(x) {
  coef(lm(x[,2] ~ x[,1]))[2]
}

umol_30_rate <- rollapply(data = df_slope,
                                   width = 20,
                                   FUN = lm_slope,
                                   by.column = FALSE) * 3600 * 24

plot(umol_30_rate)
```

## 2024-08-20 Chamber experiment: Sed only, open crate

Chamber label experiment in sed-only, open sided crate.

* Installed chamber. Sed-surface 10cm, chamber-surface 5cm.
* Started pump, purge bubbles. Large initial bubble below screen.
* Inject 5ml of argon-saturated 20 mmol/L 15NO3- solution.

Power outage at 8:30 AM. Past peak. Secondary peak of mass 46. 

```{r}
label_start <-  as.POSIXct("2024-08-19 13:40:00")

df <- read_gems("data/GEMS_tests/SerialCapture/GEMS_2024-08-20_chamber_sed_open.txt") |> 
  mutate(experiment = case_when(timestamp < label_start ~ "control",
                                TRUE ~ "label"))

plot_gems(df)

df |> 
  write_csv("data/GEMS_tests/SerialCapture/GEMS_2024-08-20_chamber_sed_open_proc.csv")
```


Calculate slope and rate using Scott's method.

Give each measurement loop an average timestamp and pivot to wide format.


```{r}
df_ts <- df %>% 
  mutate(cycle = cumsum(mass == 18)) %>% 
  group_by(cycle) %>% 
  mutate(cycle_ts = mean(timestamp)) %>% 
  ungroup()

df_wide <- df_ts %>% 
  select(timestamp = cycle_ts, mass, pressure, experiment) %>% 
  pivot_wider(names_from = mass, names_prefix = "mass_",
              values_from = pressure)
```

Natural nitrogen abundance and saturation concentration

```{r}
frac_abd_15 <- 0.00367
frac_abd_14 <- 1 - frac_abd_15
nit_sat_umol <- 600 # calculate for lab temp and pressure
```

Plot to find rough background values (should do this programmatically)

```{r}
df_wide %>% 
  select(timestamp, mass_28, mass_29, mass_30, mass_40) %>% 
  dygraph() %>% 
  dyOptions(logscale = TRUE)
```

Backgrounds from plot

```{r}
bg_29 <- 1E-11
bg_30 <- 5.3E-12
```

Normalized partial pressures and molar concentrations

```{r}
df_proc <- df_wide %>% 
  mutate(et = timestamp - timestamp[1],
         mass_28_18 = mass_28 / mass_18,
         mass_29_18 = mass_29 / mass_18,
         mass_30_18 = mass_30 / mass_18,
         mass_28_40 = mass_28 / mass_40,
         mass_29_40 = mass_29 / mass_40,
         mass_30_40 = mass_30 / mass_40,
         mass_40_28 = (mass_40 / mass_28), # include mass 40 bg subtraction?
         mass_29_28 = ( mass_29 - bg_29 ) / mass_28,
         mass_30_28 = ( mass_30 - bg_30 ) / mass_28,
         # molar concentration based on nitrogen saturation
         umol_29 = mass_29_28 * nit_sat_umol,
         umol_30 = mass_30_28 * nit_sat_umol)
```

Nitrogen normalized argon. Use slope to calculate diffusive loss.

```{r}
df_proc %>% 
  select(et, mass_40_28) %>% 
  mutate(et = as.integer(et)) %>% 
  dygraph() %>% 
  dyOptions(logscale = TRUE)
```

```{r}
ar <- df_proc %>% 
  select(et, mass_40_28) %>% 
  filter(et > 2000,
         et < 6000)
```

```{r}
ar_loss_rate <- coef(lm(mass_40_28 ~ et, data = ar))[2] * 3600 * 24
```

pick time with even slope

```{r}
df_proc %>% 
  select(et, umol_30) %>% 
  mutate(et = as.integer(et)) %>% 
  dygraph() 
```

Use least-squares linear regression to get slope of mass 30 production

```{r}
sdf <- df_proc %>% 
  select(et, umol_30) %>% 
  filter(et > 2000,
         et < 6000)
```

Slope (umol/s) of linear fit for inital linear range. Convert to umol/day

```{r}
umol_30_rate <- coef(lm(umol_30 ~ et, data = sdf))[2] * 3600 * 24
```

Initial rate of mass 30 production: `umol_30_rate` umol/day.

Rolling slope for instantaneous rate

```{r}
library(zoo)

df_slope <- df_proc %>% 
  select(et, umol_30) %>% 
  mutate(et = as.integer(et)) %>% 
  filter(et > 10000)

lm_slope <- function(x) {
  coef(lm(x[,2] ~ x[,1]))[2]
}

umol_30_rate <- rollapply(data = df_slope,
                                   width = 20,
                                   FUN = lm_slope,
                                   by.column = FALSE) * 3600 * 24

plot(umol_30_rate)
```

try with approxfun